---
title: "Explore ODA agreements targeting climate change mitigation"
format:
  html:
    code-fold: true
editor: source
execute: 
  echo: true
  warning: false
  error: false
---

## Load packages and data

Lets start by loading the packages used in this document.

Norwegian development aid statistics from 1960 onwards are available to download from [Access to microdata](https://resultater.norad.no/microdata). Select the the last ten years (2011-2021). You can also automatically download a csv file with microdata for the ten years 2011-2022 using [this link](https://resultater.norad.no/api/microdata?from_year=2011&to_year=2021&main_region_code=&country_iso_code=&agreement_partner_group_sid=&agreement_partner_sid=&target_area_code=&dac_main_sector_code=&dac_sub_sector_code=&chapter_code=&format=csv&language=en). Save the csv file in the *data* folder in the project directory and the file should be named *Norad-Norwegian_development_assistance*.

Let's import the csv file and only keep data for the years 2013-2018.

```{r}
#| label: packages-data

library(tidyverse)
library(here)
library(tidytext)

df_oda <- read_csv(here("data", "Norad-Norwegian_development_assistance.csv"))

df_oda <- df_oda |> 
  filter(year %in% c(2013:2017))
```

## Explore data

About ten percent of all agreement observations in the data are registered with the climate change adaptation marker as a main objective or significant objective. The rest is registered withouth such objectives.

Shifting focus to disbursed amounts, about 11 percent of disbursed amounts are to agreements with climate change mitigation as a main objective, and 3 percent to agreements with significant objective. 86 percent of disbursements are to agreements without such an objective.

```{r}
#| label: explore-mitigation
#| eval: false

# # Summarising number of observations and amounts targeting mitigation
# tbl_count <- df_oda |> 
#   group_by(policy_marker_climate_change_mitigation) |> 
#   summarise(
#     n = n(),
#     amount_nok_mill = sum(disbursements_1000_nok / 1000)) |>
#   mutate(
#     n_prop = n / sum(n),
#     amount_prop = amount_nok_mill / sum(amount_nok_mill)) |> 
#   ungroup()
# 
# knitr::kable(tbl_count)

```

Let's have a look at the most frequently used words in title and description of agreements with and without mitigation objective. "redd", "climate", "energy", "forest" and "sustainable" are among the most frequently used words in agreements targeting mitigation.

```{r explore-text}

# Making a binary variable "mitigation" and character variable "title_desc"
# Facet plot of frequently used words
df_oda |>
  mutate(mitigation = if_else(policy_marker_climate_change_mitigation == 0, "Not mitigation", "Mitigation")) |> 
  mutate(title_desc = paste0(agreement_title, ". ", description_of_agreement)) |> 
  unnest_tokens(word, title_desc) |>
  anti_join(get_stopwords()) |>
  count(mitigation, word, sort = TRUE) |>
  group_by(mitigation) |>
  slice_max(n, n = 15) |>
  ungroup() |>
  mutate(word = reorder_within(word, n, mitigation)) |>
  ggplot(aes(n, word, fill = mitigation)) +
  geom_col(show.legend = FALSE) +
  scale_y_reordered() +
  facet_wrap(~mitigation, scales = "free") +
  labs(x = "Word frequency",
       y = NULL) +
  theme_minimal()

```
