---
title: "Exploring text data in development aid agreements targeting climate change mitigation"
format:
  html:
    code-fold: true
editor: source
execute: 
  echo: true
  warning: false
  error: false
---

## Load packages and data

Norwegian development aid statistics (microdata) is publicly available to download [aidresults.no/microdata](https://resultater.norad.no/microdata). We select last five years (2013-2017). These data can also be downloaded by using [this link](https://resultater.norad.no/api/microdata?from_year=2011&to_year=2017&main_region_code=&country_iso_code=&agreement_partner_group_sid=&agreement_partner_sid=&target_area_code=&dac_main_sector_code=&dac_sub_sector_code=&chapter_code=&format=csv&language=en). The csv file is saved in the the *data* folder in the project directory and named *Norad-Norwegian_development_assistance.csv*.

Let's import the csv file. We use the data for the years 2013-2017.

```{r}
#| label: packages-data

library(tidyverse)
library(here)
library(tidytext)

df_oda <- read_csv(here("data", "Norad-Norwegian_development_assistance.csv"))

df_oda <- df_oda |> 
  filter(year %in% c(2013:2017))
```

## Explore data

About ten percent of all agreement observations in the data are registered with the climate change adaptation marker as a main objective or significant objective. The rest is registered withouth such objectives.

Shifting focus to disbursed amounts, about 11 percent of disbursed amounts are to agreements with climate change mitigation as a main objective, and 3 percent to agreements with significant objective. 86 percent of disbursements are to agreements without such an objective.

```{r}
#| label: explore-mitigation
#| eval: false

# # Summarising number of observations and amounts targeting mitigation
# tbl_count <- df_oda |> 
#   group_by(policy_marker_climate_change_mitigation) |> 
#   summarise(
#     n = n(),
#     amount_nok_mill = sum(disbursements_1000_nok / 1000)) |>
#   mutate(
#     n_prop = n / sum(n),
#     amount_prop = amount_nok_mill / sum(amount_nok_mill)) |> 
#   ungroup()
# 
# knitr::kable(tbl_count)

```

Let's have a look at the most frequently used words in title and description of agreements with and without mitigation objective. "redd", "climate", "energy", "forest" and "sustainable" are among the most frequently used words in agreements targeting mitigation.

```{r explore-text}

# Making a binary variable "mitigation" and character variable "title_desc"
# Facet plot of frequently used words
df_oda |>
  mutate(mitigation = if_else(policy_marker_climate_change_mitigation == 0, "Not mitigation", "Mitigation")) |> 
  mutate(title_desc = paste0(agreement_title, ". ", description_of_agreement)) |> 
  unnest_tokens(word, title_desc) |>
  anti_join(get_stopwords()) |>
  count(mitigation, word, sort = TRUE) |>
  group_by(mitigation) |>
  slice_max(n, n = 15) |>
  ungroup() |>
  mutate(word = reorder_within(word, n, mitigation)) |>
  ggplot(aes(n, word, fill = mitigation)) +
  geom_col(show.legend = FALSE) +
  scale_y_reordered() +
  facet_wrap(~mitigation, scales = "free") +
  labs(x = "Word frequency",
       y = NULL) +
  theme_minimal()

```
