---
title: "Algoritme for å identifisere avtaler relatert til utslippsreduksjon"
author: "Einar Tornes"
format: html
editor: source
execute: 
  echo: true
  warning: false
  error: false
---

```{r}
#| label: load-packages

library(tidyverse)
library(tidymodels)
library(stopwords)
library(tidytext)
library(stringr)
library(noradstats)
library(here)
library(janitor)
library(knitr)

```

Let's build a model for identifying ODA agreements targeting the mitigation of climate change.

## Load data

Importing ten years of Statsys data.

```{r}
#| label: load-data

# Load statsys data for the ten last years
df_statsys <- noradstats::read_aiddata(here("data", "statsys_ten.csv")) |> 
  janitor::clean_names()

# Exclude non-ODA agreements, exclude the frame agreement level and chose years. Sorting levels.
df_oda <- df_statsys |> 
  filter(type_of_flow == "ODA",
         type_of_agreement != "Rammeavtale",
         year %in% c(2013:2017)) |> 
  mutate(pm_climate_change_mitigation = fct_relevel(pm_climate_change_mitigation, "None", after = Inf))

```

## Explore data

About ten percent of all agreement observations in the data are registered with the climate change adaptation marker as a main objective or significant objective. The rest is registered withouth such objectives.

Shifting focus to disbursed amounts, about 11 percent of disbursed amounts are to agreements with climate change mitigation as a main objective, and 3 percent to agreements with significant objective. 86 percent of disbursements are to agreements without such an objective.

```{r}
#| label: explore-mitigation

# Summarising number of observations and amounts targeting mitigation
tbl_count <- df_oda |> 
  group_by(pm_climate_change_mitigation) |> 
  summarise(
    n = n(),
    amount_nok_mill = sum(disbursed_mill_nok)) |>
  mutate(
    n_pst = n / sum(n),
    amount_pst = amount_nok_mill / sum(amount_nok_mill)) |> 
  ungroup()

knitr::kable(tbl_count)

```

Let's have a look at the most frequently used words in title and description of agreements with and without mitigation objective. "redd", "climate", "energy", "forest" and "sustainable" are among the most frequently used words in agreements targeting mitigation.

```{r explore-text}

# Making a binary variable "mitigation" and character variable "title_desc"
df_oda <- df_oda |>
  mutate(mitigation = if_else(pm_climate_change_mitigation == "None", "Not mitigation", "Mitigation")) |> 
  mutate(title_desc = paste0(agreement_title, ". ", description_of_agreement))

# Facet plot of frequently used words
df_oda |> 
  unnest_tokens(word, title_desc) |>
  anti_join(get_stopwords()) |> 
  count(mitigation, word, sort = TRUE) |>
  group_by(mitigation) |> 
  slice_max(n, n = 15) |> 
  ungroup() |> 
  mutate(word = reorder_within(word, n, mitigation)) |> 
  
  ggplot(aes(n, word, fill = mitigation)) +
  geom_col(show.legend = FALSE) +
  scale_y_reordered() +
  facet_wrap(~mitigation, scales = "free") +
  labs(x = "Word frequency",
       y = NULL) +
  theme_minimal()

```


## Build a model

```{r}

```

